---
title: "A Storyboard on Ozone Layer"
Author: "Ashraf Alaghbari"
output: 
  flexdashboard::flex_dashboard:
    storyboard: True
    theme: journal
    source_code: embed
runtime: shiny
---


#import libraries
```{r, include=FALSE}

library(ggplot2)
library(tidyverse)
library(plotly)
library(gganimate)
library(flexdashboard)
library(ggrepel)
library(scales)
library(janitor)

```

```{r, include=FALSE}

ozone_layer<-read.csv("ozone_layer.csv")
colnames(ozone_layer)[4]<-"emissions"
b<-read.csv("b.csv")
colnames(b)[4]<-"emission"
c<-read.csv("c.csv")
colnames(c)[4]<-"emission"
chemical<-read.csv("chemicals.csv")
Vienna_Convention<-read.csv("Vienna_Convention.csv")
Ozone_hole_area<-read.csv("Ozone_hole_area.csv")
change_cfc_concentrations<-read.csv("change_cfc_concentrations.csv")
stratospheric_ozone_concentration<-read.csv("stratospheric_ozone_concentration.csv")
stratospheric_chorine_concentrations<-read.csv("stratospheric_chorine_concentrations.csv")
stratospheric_ozone_concentration_projections<-read.csv("stratospheric_ozone_concentration_projections.csv")
atmospheric_concentrations_CFC<-read.csv("atmospheric_concentrations_CFC.csv")
skin_cancer<-read.csv("skin_cancer.csv")

```

### <b>Introduction</b>

<p style="font-family: times, serif; font-size:15pt; text-align: justify">
    The emission of ozone-depleting gases impacted the earth’s ozone layer – a layer that absorbs most of the Sun’s ultraviolet radiation. Emissions were increasing rapidly until the 1980s. Since then the world has achieved rapid progress: the near-elimination of ozone-depleting substances and the trend towards recovery of the ozone layer are arguably among the most successful international environmental collaborations to date. In this entry I present the data on ozone layer depletion, signs of recovery, emissions of ozone-depleting substances, international agreement and collaboration, and the consequences of ozone layer depletion.
</p>

### <b>Emissions of ozone-depleting substances</b>



```{r, echo=FALSE}

ozone_layer<-ozone_layer%>%pivot_wider(names_from = Entity, values_from = emissions, values_fill = 0)
colnames(ozone_layer)<-colnames(ozone_layer) %>% make_clean_names()
f<-list(family="serif", size=22.4, color="#555555")
label =list(bgcolor ="white",
            bordercolor="green", 
            font=f)

colnames(ozone_layer) <- colnames(ozone_layer) %>% make_clean_names()
```



```{r Fig, echo=FALSE, fig.height=7, fig.width=15}

plot_ly() %>% add_trace(data=ozone_layer,type = 'scatter',
                                                     marker = list(color = "#3C4E66"),
                        mode = 'lines+markers', x=~year, y=~total_emissions,name="Total emissions",

 hovertemplate =paste("<b>%{x}</b><br>",
                      "Total emissions : %{y}",
                      "<extra></extra>")) %>%
  add_trace(data=ozone_layer,type = 'scatter',
            mode = 'lines+markers', x=~year, y=~natural_emissions,name="Natural emissions",
            marker = list(color = "#B13507"),
            hovertemplate =paste("<b>%{x}</b>",
                                 "<br> Natural emissions: %{y}" ,
                                 "<extra></extra>"))  %>% 
  layout(title=list(text = "<b>Ozone-depleting substance emissions in million tonnes, 1961 to 2014</b>",
                    font = f,
                    y = 1, x=0.1, hovermode = "y unified"),
         yaxis= list(title ="",range = list(0, 1500000)),
         xaxis= list(title =""), annotations = 
           list(x = 1, y = -0.07, text = "Source: Hegglin et al. (2014).", 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=0, yshift=0,
                font=list(size=12, color="black"))) %>%
    style(hoverlabel= label)
```

---

<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -Shown in the chart is the level of natural emissions (which has been approximately consistent over
this period), and total emissions which is the sum of natural and man-made emissions. Here we see a
clear growth-peak-reduction trend in ozone-depleting emissions, with a rapid rise in emissions (increasing more than three-fold) from 1960 through to the late 1980s, followed by a similarly fast reduction. By 2010, emissions had returned to 1960 levels.
</p>


### <b>Consumption of ozone-depleting substances</b>


```{r Fig1, echo=FALSE, fig.height=7, fig.width=15}
ggplot(b, aes(Year, emission))+
  geom_line(size=1, color="#3C4E66")+
  scale_y_continuous(breaks = seq(0, 150, by = 10),labels = unit_format(unit = "%"))+
  scale_x_continuous(breaks = seq(1986, 2018, by = 4))+
  theme(axis.text.x=element_text(face="bold",size=12))+
  theme(axis.text.y=element_text(face="bold",size=12))+
  geom_point()+
  labs(title = "Change in the consumption of ozone-depleting substances, world", 
    subtitle = "Global consumption of ozone-depleting substances (ODS),
measured relative to the year 1986 (where consumption in 1986 is equal to 100).", 
       caption = "Source: European Environment Agency (EEA)")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+theme(axis.title=element_blank())


```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -In the chart we see the magnitude of global decline in ODS consumption since 1986. This data measures the indexed consumption of ODS to the 1986 (i.e. levels in 1986 are equal to 100). Following a 28% rise in 1987, we see a rapid consumption in global emissions. Consumption fell by more than 60% by 1995; 80% by 2000; and by 99.7% percent by 2018.
<p>

### <b>Consumption of ozone-depleting substances by country</b>
```{r}
renderPlot({
  a<-c[,c("Year",as.character(input$select))]
  names(a)<-c("year","country")
  ggplot(a,aes(year, country))+ geom_line()+
  scale_x_continuous(breaks = seq(1986, 2018, by = 2))+
  scale_y_continuous(label=label_number(suffix = " tonnes"))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_line(size=1)+
  labs(title = "Consumption of Ozone-Depleting Substances, 1989 to 2014", 
       subtitle = "Consumption of all ozone-depleting substances (ODS). ODS consumption is
measured units of ODS tonnes, which is the amount of ODS consumed, multiplied by their respective
ozone depleting potential value. Data for individual parties to the European Union (EU) are not
shown since party obligations are collective", 
       caption = "Source: European Environment Agency (EEA)")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
  theme(axis.title=element_blank())+
  guides(color=F, size=F, fill=F)
  
})
```

***
```{r}
c<-pivot_wider(c[,-2], names_from = Entity, values_from = emission, values_fill = 0) %>% arrange(Year)
list<-as.list(colnames(c))[-1]

selectInput("select", h3("Country"),
            choices=list, selected=list[205])
```
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -In the chart we see the quantity of ODS consumption by country. This is measured in tonnes of ozone-depleting substances all weighted relative to their depleting potential.
<p>
### <b>Consumption of ozone-depleting substances by substance</b>

```{r}
renderPlot({
  z<-v[,c("Year",as.character(input$radio))]
  names(z)<-c("year","depleted_substance")

ggplot(z,aes(year, depleted_substance))+ geom_area(aes(fill="navy"))+
  scale_x_continuous(breaks = seq(1989, 2014, by = 2))+
  scale_y_continuous(labels = unit_format(unit = "M tonnes", scale = 1e-6))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_line(size=1)+
  labs(title = "Ozone-depleting substance consumption, World", 
       subtitle = "Annual consumption of ozone-depleting substances (ODS). ODS consumption is measured units of ODS tonnes, which
is the amount of ODS consumed, multiplied by their respective ozone depleting potential value.", 
       caption = "Source: UN Environment Programme")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
  theme(axis.title=element_blank())+
  guides(color=F, size=F, fill=F)

})
```

***
```{r}

v<-chemical[,c(-1,-2)]
v[is.na(v)] <- 0
v<-aggregate(v[,-1], by=list(Year=v$Year), FUN=sum)
v<-v %>% rename("Methyl Bromide"="Methyl.Bromide",
             "Carbon Tetrachloride"="Carbon.Tetrachloride",
             "Hydrochlorofluorocarbons HCFCs"="Hydrochlorofluorocarbons..HCFCs.",
             "Chlorofluorocarbons CFCs"="Chlorofluorocarbons..CFCs.",
             "Methyl Chloroform"="Methyl.Chloroform")
lis<-as.list(colnames(v))[-1]
radioButtons("radio", h3("Ozone-depleting substance"),
            choices=lis, selected=lis[1])

```
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
     -At the global level we see the trend of declining consumption — as established above — since 1989. However, it’s also interesting to note the relative decline and change in the quantity of individual substances. Throughout the 1990s and first half of the 2000s, chlorofluorocarbons (CFCs) dominated global consumption (accounting for 60 percent, reducing to 50 percent). However, through the 2000s we have seen a rising dominance of hydrochlorofluorocarbons (HCFCs); in 2014 HCFCs accounted for 94 percent of global consumption.<br><br>-This transition reflects the substitution of CFCs with HCFCs in order to reducing total ozone depletion.<br><br>-Chlorofluorocarbons (CFCs) have almost been completely phased out, declining from over 2,700,000 tonnes in 1989 to 156 tonnes in 2014.
<p>
### <b>International agreements for the Protection of the Ozone Layer</b>

```{r}
renderPlot({
  agreement<-Vienna_Convention[Vienna_Convention$international_agreement==input$select_agreement,
                               c("year","international_agreement","number_of_parties")]
ggplot(agreement,aes(year, number_of_parties))+ geom_line(size=1)+
  scale_x_continuous(breaks = seq(1971, 2015, by = 4))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  labs(title = "Number of parties in multilateral environmental agreements", 
       subtitle = "Total number of global parties signed on to multilateral agreements
designed to address trans-boundary environmental issues.", 
       caption = "Source: United Nations Conference on Trade and Development (UNCTAD)")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
  theme(axis.title=element_blank())+
  guides(color=F, size=F, fill=F)
})

```

***
```{r, include=F}

colnames(Vienna_Convention)<-gsub(".", " ", colnames(Vienna_Convention), fixed=TRUE)
Vienna_Convention<-pivot_longer(Vienna_Convention,4:16 ,names_to = "international agreement",
                values_to="number of parties")
library(janitor)
colnames(Vienna_Convention)<-colnames(Vienna_Convention) %>% make_clean_names()
```

```{r}
list_agreements<-as.list(unique(Vienna_Convention$international_agreement))
selectInput("select_agreement", h3("International agreement"),
            choices=list_agreements, selected=list_agreements[1])
```
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -The rapid decline in emissions of ozone-depleting substances shown on the left was driven by international agreement to phase out their production. In 1985 the Vienna Convention for the Protection of the Ozone Layer was adopted and entered into force in 1988.<br><br>-In the chart we see the evolution of global parties signing on to the Vienna Convention. In its first year (1988) there were only 29 parties signed to the agreement. This rapidly increased in the years to follow, reaching 174 parties by 2000. 
<p>
### <b>Ozone layer depletion: Stratospheric ozone concentration</b>

```{r, echo=F}

colnames(stratospheric_ozone_concentration)<-colnames(stratospheric_ozone_concentration)%>%
  make_clean_names()

stratospheric_ozone_concentration<-pivot_longer(stratospheric_ozone_concentration,4:5 ,names_to = "stat",
                                values_to="concentration")
stratospheric_ozone_concentration$stat<-as.factor(stratospheric_ozone_concentration$stat)
levels(stratospheric_ozone_concentration$stat)<-list("mean daily concentration"="mean_daily_concentration_nasa",
                                                     "minimum daily concentration"="minimum_daily_concentration_nasa")
```


```{r, echo=F}
a<-stratospheric_ozone_concentration %>%
  mutate(label = if_else(year == max(year), as.character(stat), NA_character_)) 
```



```{r Fig2, echo=FALSE, fig.height=7, fig.width=15}

ggplot(a,aes(year, concentration, color=stat))+scale_y_continuous(label=label_number(suffix = " DU"))+ scale_x_continuous(breaks = seq(1979, 2020, by = 4))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_line(size=1)+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE,direction = c("y"))+
  labs(title = "Stratospheric ozone concentration", 
    subtitle = "Stratospheric ozone concentration in the Southern Hemisphere,measured in
Dobson Units   (DU). One Dobson Unit is the number of molecules of ozone that would be required to create a layer of pure ozone 0.01 millimeters thick at a temperature of 0 degrees
Celsius and a pressure of 1 atmosphere.", 
       caption = "Source: NASA Ozone Watch")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+theme(axis.title=element_blank())+
  guides(color=F, size=F)+scale_color_manual(values=c("#B13507","#3C4E66"))


```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -What impact has man-made ODS emissions had on stratospheric ozone concentrations?<br><br>-In the chart we see average stratospheric ozone concentrations in the Southern Hemisphere (where ozone depletion has been most severe) from 1979 to 2020.<br><br>-We see that since 1979 through to the early 1990s, stratospheric ozone concentrations in the South Hemisphere fell to the concerning ‘ozone hole’ level of 100 DU. For several decades since the 1990s, concentrations have continued to approximate around (or below) 100 DU. Over the last few years since 2010, however, ozone concentrations have started to slowly recover.
<p>

### <b>Ozone layer depletion: Ozone hole area</b>

```{r}


colnames(Ozone_hole_area)<-colnames(Ozone_hole_area)%>%
  make_clean_names()

Ozone_hole_area<-pivot_longer(Ozone_hole_area,4:5 ,names_to = "stat",
                                                values_to="area")
Ozone_hole_area$stat<-as.factor(Ozone_hole_area$stat)
levels(Ozone_hole_area$stat)<-list("maximum ozone hole area nasa"="maximum_ozone_hole_area_nasa",
                                                     "mean ozone hole area"="mean_ozone_hole_area")
Ozone_hole_area<-Ozone_hole_area %>%
  mutate(label = if_else(year == max(year), as.character(stat), NA_character_))
```

```{r Fig3, echo=FALSE, fig.height=7, fig.width=15}
 
ggplot(Ozone_hole_area,aes(year, area, color=stat))+
  scale_y_continuous(labels = unit_format(unit = "M km²", scale = 1e-6))+
  scale_x_continuous(breaks = seq(1979, 2020, by = 4))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_line(size=1)+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE,direction = c("y"))+
  labs(title = "Antarctic ozone hole area", 
       subtitle = "Annual maximum and mean Antarctic stratospheric ozone hole area,
resultant from the emission of ozone-depleting substances.", 
       caption = "Source: NASA Ozone Watch")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+theme(axis.title=element_blank())+
  guides(color=F, size=F)+scale_color_manual(values=c("#B13507","#3C4E66"))

```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -Since 1979 we see a distinct increase in the Antarctic ozone hole area, reaching a maximum of 30 million km2 in the early 2000s. However, since the late 1990s, the ozone hole area had approximately stabilised between 20 to 25 million km2. 
<p>

### <b>When is the ozone layer expected to recover?</b>

```{r}

stratospheric_chorine_concentrations<-rename(stratospheric_chorine_concentrations,
          "chorine_concentrations"="Equivalent.stratospheric.chorine..ESC...1960...0.", "year"="Year")

stratospheric_chorine_concentrations<-stratospheric_chorine_concentrations %>%
  mutate(label = if_else(year == max(year), as.character(Entity), NA_character_))
```

```{r Fig4, echo=FALSE, fig.height=7, fig.width=15}

ggplot(stratospheric_chorine_concentrations,aes(year, chorine_concentrations, color=Entity))+
  scale_x_continuous(breaks = seq(1960, 2100, by = 10))+
  scale_y_continuous(breaks = seq(0, 60, by = 10))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_line(size=1)+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE,direction = c("y"))+
  labs(title = "Stratospheric chlorine concentrations, 1960 to 2100", 
       subtitle = "Equivalent stratospheric chloric (ESC) concentrations with projections to 2100
based on chemistry-climate models. ESC concentrations are measured relative to
levels in 1960 (1960 = 0)", 
       caption = "Source: Hegglin et al. (2014).")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
  theme(axis.title=element_blank())+
  guides(color=F, size=F)+scale_color_manual(values=c("#B13507","#3C4E66","#0073E6"))

```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -The data presented is measured relative to concentrations in 1960 (where 1960 is equal to 0). ODS can have a significant lifetime in the atmosphere, for some between 50 and 100 years on average. This means that despite reductions in ODS emissions (and eventually complete phase-out of these substances), equivalent stratospheric chlorine (ESC) concentrations are expected to remain higher than 1960 levels through to the end of the century. However, it’s expected that they peaked in the early 2000s and will continue to slowly decline throughout this period
<p>

### <b>When is the ozone layer expected to recover?</b>

```{r}

stratospheric_ozone_concentration_projections<-rename(stratospheric_ozone_concentration_projections,  "ozone_concentration_projections"="Ozone.concentration..1960...0.","year"="Year")
stratospheric_ozone_concentration_projections<-stratospheric_ozone_concentration_projections %>%
  mutate(label = if_else(year == max(year), as.character(Entity), NA_character_))

```

```{r Fig5, echo=FALSE, fig.height=7, fig.width=15}

ggplot(stratospheric_ozone_concentration_projections,aes(year, ozone_concentration_projections, color=Entity))+
  geom_line(aes(alpha=factor(Entity)), size=1)+
  scale_x_continuous(breaks = seq(1960, 2100, by = 10))+
  scale_y_continuous(breaks = seq(-140, 40, by = 20))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE,direction = c("y"))+
  labs(title = "Stratospheric ozone concentration projections, 1960 to 2100", 
       subtitle = "Stratospheric ozone concentrations with projections to 2100 based on
chemistry-climate models. Ozone concentrations are measured relative
to levels in 1960 (1960 = 0), and measured as the global average,and regional
average. Figures represent the mean across a number of model runs; model projections
have notable uncertainty around such average trends.", 
       caption = "Source: Hegglin et al. (2014).")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
  theme(axis.title=element_blank())+
  theme(legend.position = "none")+scale_color_manual(values=c("#B13507","#3C4E66","#0073E6"))+
  geom_hline(aes(yintercept = 0), alpha=0.2,size=0.6)+scale_alpha_manual(values=c(0.2,0.2,1))+
  geom_point()

```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -In the chart, we see stratospheric ozone concentrations relative to 1960 (1960 = 0). As a global average concentration, it’s expected that ozone levels will return to their 1960 levels around mid-century. Antarctica, where ozone depletion has been most severe due to very low temperatures is expected to recover much more slowly. It’s projected that Antarctic ozone concentrations will only begin to approach 1960 levels by the end of the century.
<p>

### <b>How was an increase in emissions detected?</b>

```{r}

colnames(change_cfc_concentrations)<-colnames(change_cfc_concentrations)%>%
  make_clean_names()

change_cfc_concentrations<-pivot_longer(change_cfc_concentrations,4:5 ,names_to = "stat",
                              values_to="concentration")

change_cfc_concentrations$stat<-as.factor(change_cfc_concentrations$stat)
levels(change_cfc_concentrations$stat)<-list("Actual atmospheric CFC-11 concentrations"="cfc_11_measurement",
                                   "Expected/projected atmospheric CFC-11 concentrations"="wmo_2014")


change_cfc_concentrations<-change_cfc_concentrations %>%
  mutate(label = if_else(year == max(year), as.character(stat), NA_character_))
change_cfc_concentrations$concentration[is.na(change_cfc_concentrations$concentration)]<-0

```

```{r Fig6, echo=FALSE, fig.height=7, fig.width=15}

ggplot(change_cfc_concentrations,aes(year, concentration))+
  geom_line(aes( group = stat,color=stat, linetype = stat),size=1)+
  theme(legend.position = "none")+ 
  scale_x_continuous(breaks = seq(1995, 2018, by = 1))+
  scale_y_continuous(breaks = seq(-1.2, 0, by = 0.2),label=label_number(suffix = "%"))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE,direction = c("y"))+
  labs(title = "Annual change in atmospheric CFC-11 concentrations", 
       subtitle = "Annual change in atmospheric trichlorofluoromethane (CFC-11) concentrations, given as actual measured concentration
change (solid line), and expected/projected rate of change based on reported emissions of parties to the Montreal
Protocol (dashed line). Since 2012, concentrations have fallen at a lower rate than expected, suggesting unreported
sources of CFC-11 emissions.", 
       caption = "Source: Montzka et al. (2018)")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
  theme(axis.title=element_blank())+scale_color_manual(values=c("#555555","#555555"))+geom_point()

```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -In the chart we see the annual change (in percent) of measured concentrations of CFC-11 (shown as the solid line). Also shown is the 2014 projections/expectations of the rate of change in concentration based on reported emissions to the Montreal Protocol (shown as the dashed line). As we see, actual and expected concentration changes map closely over the period up to 2011. Since 2012, however, the annual rate of decline in concentrations has fallen (almost halved from -0.8 percent to -0.4 percent per year). 
<p>

### <b>Will recent emissions hinder progress on ozone recovery?</b>

```{r}


colnames(atmospheric_concentrations_CFC)<-colnames(atmospheric_concentrations_CFC)%>%
  make_clean_names()

atmospheric_concentrations_CFC<-pivot_longer(atmospheric_concentrations_CFC,4:6 ,names_to = "stat",
                                        values_to="concentration")

atmospheric_concentrations_CFC$stat<-as.factor(atmospheric_concentrations_CFC$stat)
levels(atmospheric_concentrations_CFC$stat)<-list("northern hemisphere"="northern_hemisphere",
                                             "southern hemisphere"="southern_hemisphere",
                                             "WMO 2014 projected"="wmo_2014")

atmospheric_concentrations_CFC<-atmospheric_concentrations_CFC %>% filter(year >= 1995 & year <= 2017)
atmospheric_concentrations_CFC<-atmospheric_concentrations_CFC %>%
  mutate(label = if_else(year == max(year), as.character(stat), NA_character_))

```

```{r Fig7, echo=FALSE, fig.height=7, fig.width=15}

ggplot(atmospheric_concentrations_CFC,aes(year, concentration))+
  geom_line(aes( group = stat, color=stat, linetype = stat),size=1)+
  scale_linetype_manual(values=c("solid","solid","dashed"))+
  theme(legend.position = "none")+ 
  scale_x_continuous(breaks = seq(1995, 2018, by = 1))+
  scale_y_continuous(breaks = seq(200, 260, by = 10),label=label_number(suffix = " ppt"))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE,direction = c("both"))+
  labs(title = "Measured and expected atmospheric concentrations of CFC-11", 
       subtitle = "Measured concentrations of trichlorofluoromethane (CFC-11) gas in
the Northern and Southern hemispheres, measured relative to the expected/projected
concentrations based on emissions reported by countries to the Montreal Protocol.
This is measured in mole fractions (ppt) of CFC-11. In recent years (since 2012),
concentrations have not declined at the projected rate,
suggesting unreported sources of CFC-11 emissions.", 
       caption = "Source: Montzka et al. (2018)")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
  theme(axis.title=element_blank())+scale_color_manual(values=c("#0073E7","#B13507","grey"))+geom_point()

```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -The long-term impact of emissions for the ozone layer will depend on how long continued emissions of CFC-11 persist.<br><br>-Despite recent emissions, total concentrations continue to fall but at a notably slower rate than expected
<p>

### <b>Impacts of ozone depletion on skin cancer risk</b>

```{r}

colnames(skin_cancer)<-colnames(skin_cancer)%>%
  make_clean_names()

skin_cancer<-pivot_longer(skin_cancer,4:6 ,names_to = "stat",
                                             values_to="excess_skin_cancer_cases")
skin_cancer<-skin_cancer %>% filter(entity=="United States")
skin_cancer$stat<-as.factor(skin_cancer$stat)
levels(skin_cancer$stat)<-list("Copenhagen Amendments"="excess_skin_cancer_cases_per_million_copenhagen_amendments",
                                                  "Montreal Protocol"="excess_skin_cancer_cases_per_million_montreal_protocol",
                                                  "No restrictions"="excess_skin_cancer_cases_per_million_no_restrictions")
skin_cancer<-skin_cancer %>%
  mutate(label = if_else(year == max(year), as.character(stat), NA_character_))

```

```{r Fig8, echo=FALSE, fig.height=7, fig.width=15}

ggplot(skin_cancer,aes(year, excess_skin_cancer_cases, color=stat))+
  geom_line(size=1)+
  theme(legend.position = "none")+ 
  scale_x_continuous(breaks = seq(2000, 2100, by = 20))+
  scale_y_continuous(breaks = seq(0, 6000, by = 1000))+
  theme(axis.text.x=element_text(color="#333333",size=14.4))+
  theme(axis.text.y=element_text(color="#333333",size=14.4))+
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE,direction = c("y"))+
  labs(title = "Excess skin cancer cases due to ozone depletion, United States, 2000 to 2100", 
       subtitle = "Historic estimates of the number of excess skin cancer cases,
measured per million people, resultant from stratospheric ozone depletion
(which increases exposure to UV radiation). Estimated under scenario
of no restrictions on ozone-depleting substances, the initial Montreal
Protocol and Copenhagen Amendment, which tightened regulation.", 
       caption = "Source: Slaper et al. (1996)")+
  theme(plot.title = element_text(hjust = 0, vjust = 1,color="#555555",
                                  size=22.4,family="serif"))+
theme(axis.title=element_blank())+
  scale_color_manual(values=c("#0073E7","#B13507","grey"))+geom_point()

```

***
<p style="font-family: times, serif; font-size:13pt; text-align: justify">
    -The study estimates that by 2030, two million cases worldwide will be avoided per year as a result of ODS reduction from the Montreal Protocol and its later revisions.
This is shown for three scenarios: the assumption of no protocol/restrictions on ODS consumption; the initial Montreal Protocol treaty, and the later Copenhagen Amendment which increased ambition for ODS reduction.
<p>

### <b>Conclusion</b>
<p style="font-family: times, serif; font-size:15pt; text-align: justify">
    -Global emissions of ozone-depleting substances have declined by more than 99% since 1986 (the year before international action was agreed).
<p>

<p style="font-family: times, serif; font-size:15pt; text-align: justify">
    -The Montreal Protocol (and later amendments) was adopted in 1987 — since then all countries have signed on to the agreement, allowing for the dramatic decline in global ozone-depleting emissions.
<p>

<p style="font-family: times, serif; font-size:15pt; text-align: justify">
    -Ozone layer thickness declined, and the Antarctic ozone hole grew substantially from the 1980s through to the early 2000s. Through the first decade this trend largely stabilized and we now see initial signs of recovery.
<p>

<p style="font-family: times, serif; font-size:15pt; text-align: justify">
    -It will take decades for the ozone layer to recover.
<p>

<p style="font-family: times, serif; font-size:15pt; text-align: justify">
    -Ozone layer depletion increases the amount of ultraviolet (UV) irradiation that reaches Earth’s surface; this can increase the risk of skin cancer, particularly at higher latitudes.
<p>

<p style="font-family: times, serif; font-size:15pt; text-align: justify">
    -The global shift away from ozone-depleting substances also had co-benefits on the reduction of greenhouse gas emissions.
<p>


### <b>Data sources</b>

<h1 style="font-size:2vw"><b>NASA Ozone Watch</b></h1>
<b>Data:</b> Concentration of atmospheric ozone and ozone hole area
<br>
<b>Geographical coverage:</b> Global
<br>
<b>Time span:</b> 1979-2017
<br>
<b>Available at:</b><a href="https://ozonewatch.gsfc.nasa.gov/meteorology/annual_data.html"> NASA Ozone Hole Watch</a>

<h1 style="font-size:2vw"><b>NASA Monthly Ozone Maps</b></h1>
<b>Data:</b> Monthly satellite maps of Antarctic ozone
<br>
<b>Geographical coverage:</b> Antarctic pole
<br>
<b>Time span:</b> 1979-present (monthly)
<br>
<b>Available at:</b><a href="https://ozonewatch.gsfc.nasa.gov/monthly/SH.html"> NASA Ozone Hole Watch Satellite Imaging</a>


<h1 style="font-size:2vw"><b>Ozone Secretariat Scientific Assessment Panel</b></h1>
<b>Data:</b> Synthesis reports on ‘Twenty Questions and Answers About the Ozone Layer
<br>
<b>Geographical coverage:</b> Antarctic pole
<br>
<b>Time span:</b> Published every four years (2006; 2010; 2014; 2018)
<br>
<b>Available at:</b><a href="https://web.archive.org/web/20180622084056/http://ozone.unep.org/en/assessment-panels/scientific-assessment-panel"> Ozone Secretariat Scientific Assessment Panel</a>





